# Data reorganisation and ANOVA (1.3)

## Load the data

Each row of `Results.txt` is one movement: group, participant (`SUBJ`), trial, orientation
(`ORI`), direction (`DIR`), index of difficulty (`ID`), repetition (`REP`), number of velocity
peaks (`NbVelPeaks`, i.e. the number of sub-movements) and movement time in seconds.

```{r}
#| label: setup
#| include: true
library(tidyverse)  # data wrangling and ggplot2
library(ez)         # mixed-design ANOVA (ezANOVA, ezPlot, ezStats)
library(here)       # paths relative to the project root
theme_set(theme_minimal(base_size = 11))
```

```{r}
#| label: load
raw <- read.table(here("data", "Results.txt"), header = TRUE,
                  stringsAsFactors = TRUE)
glimpse(raw)
```
## Check for missing data

A mixed ANOVA needs every participant to have a value in every cell of the within-subjects
factors. We count the movements in each combination of factors.

```{r}
#| label: count-id
raw |> count(GROUP, ID)
```

The group `pp` has only 400 movements for ID3, against 760 for ID1 and ID2: many patients
did not perform ID3, the most difficult condition. We remove ID3.

```{r}
#| label: drop-id3
d <- raw |> filter(ID != "ID3") |> droplevels()
levels(d$ID)
d |> count(GROUP, ID)
```

We now check the repetitions.

```{r}
#| label: count-rep
d |> count(REP)
d |> count(GROUP, REP)
```

The third repetition is missing for the group `pp` and incomplete for the group `ap`
(440 instead of 520 movements). We remove R3.