# Data reorganisation and ANOVA (1.3)

## Load the data

Each row of `Results.txt` is one movement: group, participant (`SUBJ`), trial, orientation
(`ORI`), direction (`DIR`), index of difficulty (`ID`), repetition (`REP`), number of velocity
peaks (`NbVelPeaks`, i.e. the number of sub-movements) and movement time in seconds.

```{r}
#| label: setup
#| include: true
library(tidyverse)  # data wrangling and ggplot2
library(ez)         # mixed-design ANOVA (ezANOVA, ezPlot, ezStats)
library(here)       # paths relative to the project root
theme_set(theme_minimal(base_size = 11))
```

```{r}
#| label: load
raw <- read.table(here("data", "Results.txt"), header = TRUE,
                  stringsAsFactors = TRUE)
glimpse(raw)
```
## Check for missing data

A mixed ANOVA needs every participant to have a value in every cell of the within-subjects
factors. We count the movements in each combination of factors.

```{r}
#| label: count-id
raw |> count(GROUP, ID)
```

The group `pp` has only 400 movements for ID3, against 760 for ID1 and ID2: many patients
did not perform ID3, the most difficult condition. We remove ID3.

```{r}
#| label: drop-id3
d <- raw |> filter(ID != "ID3") |> droplevels()
levels(d$ID)
d |> count(GROUP, ID)
```

We now check the repetitions.

```{r}
#| label: count-rep
d |> count(REP)
d |> count(GROUP, REP)
```

The third repetition is missing for the group `pp` and incomplete for the group `ap`
(440 instead of 520 movements). We remove R3.

```{r}
#| label: drop-r3
d <- d |> filter(REP != "R3") |> droplevels()
d |> count(GROUP, REP)
d |> distinct(GROUP, SUBJ) |> count(GROUP, name = "count")
```

The final data set has 13 age-matched adults, 19 young adults and 19 patients
(51 participants), which matches the 48 error degrees of freedom of the article
(51 − 3 groups = 48).


## One value per participant and condition

The ANOVA must compare participants, not movements: otherwise the hundreds of movements of
one participant would be treated as independent observations (pseudo-replication). We
average the movements of each participant in each ID.

```{r}
#| label: aggregate
mt <- d |>
  group_by(GROUP, SUBJ, ID) |>
  summarise(MT = mean(MovementTime), .groups = "drop")
mt |> count(GROUP, ID)
```

## Run the ANOVA

`ezANOVA()` takes the dependent variable (`dv`), the participant identifier (`wid`), the
between-subjects factor and the within-subjects factor. We ask for type III sums of
squares because the groups have unequal sizes (13, 19 and 19 participants).

```{r}
#| label: anova
aov_mt <- ezANOVA(data = mt, dv = MT, wid = SUBJ,
                  between = GROUP, within = ID, type = 3)
aov_mt$ANOVA
aov_tab <- aov_mt$ANOVA
```

With only two levels, the within-subjects factor ID cannot violate sphericity, so `ez` does
not report Mauchly's test and no Greenhouse–Geisser correction is needed.

## Visualise the interaction

`ezPlot()` draws the cell means. Its error bars are Fisher's Least Significant Difference
(FLSD) bars: two means whose bars do not overlap differ significantly.

```{r}
#| label: fig-interaction
#| fig-cap: "Movement time as a function of ID in each group (ezPlot; error bars: Fisher's Least Significant Difference). The lines are clearly not parallel: the effect of ID is much larger for the patients (pp)."
ezPlot(data = mt, dv = MT, wid = SUBJ, between = GROUP, within = ID,
       x = ID, split = GROUP, type = 3,
       x_lab = "ID", y_lab = "Movement Time (s)")
```

`ezStats()` gives the descriptive statistics of each cell (number of participants, mean,
SD and FLSD). We keep them for the next questions.

```{r}
#| label: means-table
mt_summary <- ezStats(data = mt, dv = MT, wid = SUBJ,
                      between = GROUP, within = ID, type = 3) |>
  rename(n = N, mean = Mean, sd = SD) |>
  as_tibble()
mt_summary |> mutate(across(c(mean, sd, FLSD), \(x) round(x, 2)))
```
## Are the assumptions met?

`ezANOVA()` reports Levene's test only for purely between-subjects designs, so we run it
at each ID with `car::leveneTest()` (installed with `ez`). We check normality in each cell
with the Shapiro–Wilk test.

```{r}
#| label: assumptions
mt |>                                               # equal variances between groups
  group_by(ID) |>
  group_modify(\(df, key) {
    lt <- car::leveneTest(MT ~ GROUP, data = df)
    tibble(F = lt$`F value`[1], DFn = lt$Df[1], DFd = lt$Df[2], p = lt$`Pr(>F)`[1])
  })
mt |>                                               # normality in each cell
  group_by(GROUP, ID) |>
  summarise(W = shapiro.test(MT)$statistic, p = shapiro.test(MT)$p.value,
            .groups = "drop")
```

