# Data reorganisation and ANOVA (1.3)

## Load the data

Each row of `Results.txt` is one movement: group, participant (`SUBJ`), trial, orientation
(`ORI`), direction (`DIR`), index of difficulty (`ID`), repetition (`REP`), number of velocity
peaks (`NbVelPeaks`, i.e. the number of sub-movements) and movement time in seconds.

```{r}
#| label: setup
#| include: true
library(tidyverse)  # data wrangling and ggplot2
library(ez)         # mixed-design ANOVA (ezANOVA, ezPlot, ezStats)
library(here)       # paths relative to the project root
theme_set(theme_minimal(base_size = 11))
```

```{r}
#| label: load
raw <- read.table(here("data", "Results.txt"), header = TRUE,
                  stringsAsFactors = TRUE)
glimpse(raw)
```
## Check for missing data

A mixed ANOVA needs every participant to have a value in every cell of the within-subjects
factors. We count the movements in each combination of factors.

```{r}
#| label: count-id
raw |> count(GROUP, ID)
```

The group `pp` has only 400 movements for ID3, against 760 for ID1 and ID2: many patients
did not perform ID3, the most difficult condition. We remove ID3.

```{r}
#| label: drop-id3
d <- raw |> filter(ID != "ID3") |> droplevels()
levels(d$ID)
d |> count(GROUP, ID)
```

We now check the repetitions.

```{r}
#| label: count-rep
d |> count(REP)
d |> count(GROUP, REP)
```

The third repetition is missing for the group `pp` and incomplete for the group `ap`
(440 instead of 520 movements). We remove R3.

```{r}
#| label: drop-r3
d <- d |> filter(REP != "R3") |> droplevels()
d |> count(GROUP, REP)
d |> distinct(GROUP, SUBJ) |> count(GROUP, name = "count")
```

The final data set has 13 age-matched adults, 19 young adults and 19 patients
(51 participants), which matches the 48 error degrees of freedom of the article
(51 − 3 groups = 48).


## One value per participant and condition

The ANOVA must compare participants, not movements: otherwise the hundreds of movements of
one participant would be treated as independent observations (pseudo-replication). We
average the movements of each participant in each ID.

```{r}
#| label: aggregate
mt <- d |>
  group_by(GROUP, SUBJ, ID) |>
  summarise(MT = mean(MovementTime), .groups = "drop")
mt |> count(GROUP, ID)
```

## Run the ANOVA

`ezANOVA()` takes the dependent variable (`dv`), the participant identifier (`wid`), the
between-subjects factor and the within-subjects factor. We ask for type III sums of
squares because the groups have unequal sizes (13, 19 and 19 participants).

```{r}
#| label: anova
aov_mt <- ezANOVA(data = mt, dv = MT, wid = SUBJ,
                  between = GROUP, within = ID, type = 3)
aov_mt$ANOVA
aov_tab <- aov_mt$ANOVA
```

With only two levels, the within-subjects factor ID cannot violate sphericity, so `ez` does
not report Mauchly's test and no Greenhouse–Geisser correction is needed.

## Visualise the interaction

`ezPlot()` draws the cell means. Its error bars are Fisher's Least Significant Difference
(FLSD) bars: two means whose bars do not overlap differ significantly.

```{r}
#| label: fig-interaction
#| fig-cap: "Movement time as a function of ID in each group (ezPlot; error bars: Fisher's Least Significant Difference). The lines are clearly not parallel: the effect of ID is much larger for the patients (pp)."
ezPlot(data = mt, dv = MT, wid = SUBJ, between = GROUP, within = ID,
       x = ID, split = GROUP, type = 3,
       x_lab = "ID", y_lab = "Movement Time (s)")
```

`ezStats()` gives the descriptive statistics of each cell (number of participants, mean,
SD and FLSD). We keep them for the next questions.

```{r}
#| label: means-table
mt_summary <- ezStats(data = mt, dv = MT, wid = SUBJ,
                      between = GROUP, within = ID, type = 3) |>
  rename(n = N, mean = Mean, sd = SD) |>
  as_tibble()
mt_summary |> mutate(across(c(mean, sd, FLSD), \(x) round(x, 2)))
```
## Are the assumptions met?

`ezANOVA()` reports Levene's test only for purely between-subjects designs, so we run it
at each ID with `car::leveneTest()` (installed with `ez`). We check normality in each cell
with the Shapiro–Wilk test.

```{r}
#| label: assumptions
mt |>                                               # equal variances between groups
  group_by(ID) |>
  group_modify(\(df, key) {
    lt <- car::leveneTest(MT ~ GROUP, data = df)
    tibble(F = lt$`F value`[1], DFn = lt$Df[1], DFd = lt$Df[2], p = lt$`Pr(>F)`[1])
  })
mt |>                                               # normality in each cell
  group_by(GROUP, ID) |>
  summarise(W = shapiro.test(MT)$statistic, p = shapiro.test(MT)$p.value,
            .groups = "drop")
```

The variances are not equal between groups (Levene's test, p < .001 at both IDs; the SD of the
patients is more than ten times that of the young adults), and movement time departs
from normality in two of the six cells (ID2 for the age-matched adults and for the
patients, Shapiro–Wilk p < .05). With unequal group sizes, this makes the p-values
approximate. However, the effects are so large (p < .001 for all three effects) that this
does not change the conclusion; a log transformation of MT, which reduces the skewness,
would be a reasonable robustness check.

## Result sentence {#sec-result-sentence}

```{r}
#| label: result-values
#| include: false
fmt_eff <- function(effect) {
  r <- aov_tab[aov_tab$Effect == effect, ]
  p_txt <- if (r$p < .001) "p < .001" else paste0("p = ", round(r$p, 3))
  sprintf("*F*(%d, %d) = %.2f, %s, $\\eta^2_G$ = %.2f",
          r$DFn, r$DFd, r$F, p_txt, r$ges)
}
m <- function(g, i) round(mt_summary$mean[mt_summary$GROUP == g & mt_summary$ID == i], 2)
```

Movement time increased with the index of difficulty (`r fmt_eff("ID")`) and was longer for
the patients than for both control groups (`r fmt_eff("GROUP")`). Crucially, the effect of
difficulty was not the same in all groups (GROUP × ID interaction, `r fmt_eff("GROUP:ID")`):
from ID1 to ID2, movement time increased by `r round(m("pp","ID2") - m("pp","ID1"), 2)` s in
the patients (`r m("pp","ID1")` to `r m("pp","ID2")` s), but only by
`r round(m("ap","ID2") - m("ap","ID1"), 2)` s in the age-matched adults and
`r round(m("cp","ID2") - m("cp","ID1"), 2)` s in the young adults.
