---
title: "_03-regression-graph"
format: html
editor: 
  markdown: 
    wrap: 72
---

# Run the linear regression by GROUP

To investigate the relationship between movement time and task
difficulty, we fit one linear regression for each group. The dependent
variable is movement time (`MT`) and the predictor is the index of
difficulty (`ID`).

The regression is performed on the participant-level means created in
the previous section. This avoids treating the repeated movements of one
participant as independent observations

## Prepare the data

`ID` is a categorical variable with values `ID1` and `ID2`. We create a
numeric version for the regression, where `ID1` becomes 1 and `ID2`
becomes 2.

```{r}

#| label: regression-data

mt_regression <- mt |>
  mutate(
    ID_numeric = as.numeric(sub("ID", "", ID))
  )

mt_regression
```

## Fit one regression per group

```{r}
#| label: group-regressions

regression_models <- mt_regression |>
  group_by(GROUP) |>
  group_modify(\(df, key) {
    model <- lm(MT ~ ID_numeric, data = df)

    tibble(
      intercept = unname(coef(model)[1]),
      slope = unname(coef(model)[2]),
      r_squared = summary(model)$r.squared,
      p_value = summary(model)$coefficients[2, 4]
    )
  }) |>
  ungroup()

regression_models
```

The regression equation is:

\[ MT = a + b \times ID \]

where (a) is the intercept and (b) is the slope. The slope indicates the
increase in movement time associated with a one-unit increase in task
difficulty. Therefore, the group with the largest slope is the group
most affected by increasing task difficulty.


