---
title: "_03-regression-graph"
format: html
editor: 
  markdown: 
    wrap: 72
---

# Run the linear regression by GROUP

To investigate the relationship between movement time and task
difficulty, we fit one linear regression for each group. The dependent
variable is movement time (`MT`) and the predictor is the index of
difficulty (`ID`).

The regression is performed on the participant-level means created in
the previous section. This avoids treating the repeated movements of one
participant as independent observations

## Prepare the data

`ID` is a categorical variable with values `ID1` and `ID2`. We create a
numeric version for the regression, where `ID1` becomes 1 and `ID2`
becomes 2.

```{r}

#| label: regression-data

mt_regression <- mt |>
  mutate(
    ID_numeric = as.numeric(sub("ID", "", ID))
  )

mt_regression
```

## Fit one regression per group

```{r}
#| label: group-regressions

regression_models <- mt_regression |>
  group_by(GROUP) |>
  group_modify(\(df, key) {
    model <- lm(MT ~ ID_numeric, data = df)

    tibble(
      intercept = unname(coef(model)[1]),
      slope = unname(coef(model)[2]),
      r_squared = summary(model)$r.squared,
      p_value = summary(model)$coefficients[2, 4]
    )
  }) |>
  ungroup()

regression_models
```

The regression equation is:

\[ MT = a + b \times ID \]

where (a) is the intercept and (b) is the slope. The slope indicates the
increase in movement time associated with a one-unit increase in task
difficulty. Therefore, the group with the largest slope is the group
most affected by increasing task difficulty.

```{r}
#| label: regression-equations

equations <- regression_models |>
  mutate(
    equation = sprintf(
      "MT = %.3f %s %.3f x ID",
      intercept,
      if_else(slope >= 0, "+", "-"),
      abs(slope)
    )
  )

equations |>
  select(GROUP, equation, r_squared, p_value)
```

The regression slopes can be compared with the ANOVA interaction. A
significant GROUP × ID interaction means that the effect of difficulty
differs between groups. The group with the steepest regression line
should show the largest increase in movement time from ID1 to ID2.

# Save a graph in a PDF file for publication

The following graph shows the participant-level mean movement times, the
regression line for each group, and the equation of each regression.

```{r}
#| label: equation-positions
#| include: false

equation_labels <- equations |>
  mutate(
    x = 1.05,
    y = c(3.0, 2.6, 2.2)
  )
```

```{r}
#| label: publication-graph
#| fig-cap: "Mean movement time as a function of index of difficulty, with separate linear regression lines and equations for the three groups."
#| fig-width: 8
#| fig-height: 6

publication_plot <- ggplot(
  mt_regression,
  aes(
    x = ID_numeric,
    y = MT,
    colour = GROUP,
    group = GROUP
  )
) +
  geom_point(
    alpha = 0.25,
    size = 1.8
  ) +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = FALSE,
    linewidth = 0.9
  ) +
  geom_text(
    data = equation_labels,
    aes(
      x = x,
      y = y,
      label = equation,
      colour = GROUP
    ),
    inherit.aes = FALSE,
    hjust = 0,
    show.legend = FALSE
  ) +
  geom_blank(
    data = tibble(
      ID_numeric = c(0, 3),
      MT = c(0, 6)
    ),
    aes(x = ID_numeric, y = MT),
    inherit.aes = FALSE
  ) +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("ID1", "ID2"),
    limits = c(0, 3)
  ) +
  scale_y_continuous(
    limits = c(0, 6)
  ) +
  labs(
    title = "Movement time as a function of task difficulty",
    x = "Index of difficulty (ID)",
    y = "Mean movement time (s)",
    colour = "Group"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "right"
  )

publication_plot

```

The lines are not parallel: the increase in movement time between ID1
and ID2 is larger for some groups (the stroke survivors) than for others
(the healthy age-matched and young participants). This graphical result
illustrates the significant GROUP × ID interaction found in the ANOVA.
