# Differences with the article (1.7)

## Putting our equations on the same scale as the article

In the regression section, ID was coded 1 for ID1 and 2 for ID2 (`ID_numeric`). The slopes of `regression_models` are therefore in seconds **per ID step**, and the intercepts are the movement times predicted at a step "0" that does not exist. The article fits Fitts' law on the index of difficulty in **bits** [@mottet2017; @fitts1954]. The two IDs of the task are
3.80 and 5.43 bits (x-axis of Fig 3 of the article), so one ID step is 1.63 bits.

To compare our equations with those of the article, we convert them to bits. The line goes through the same two group means; only the unit of the x-axis changes:

   - slope (s/bit) = slope (s/step) / 1.63;
   - intercept (s) = predicted MT at ID1 − slope (s/bit) × 3.80.
   

```{r}
#| label: compare-article
id_bits <- c(ID1 = 3.80, ID2 = 5.43)
step_bits <- unname(id_bits["ID2"] - id_bits["ID1"])

article <- tibble(
  GROUP = c("pp", "ap", "cp"),
  intercept_article = c(0.422, 0.622, 0.475),
  slope_article     = c(0.672, 0.118, 0.097)
)

comparison <- regression_models |>
  mutate(GROUP = as.character(GROUP),
         slope_ours     = slope / step_bits,
         intercept_ours = (intercept + slope * 1) - slope_ours * id_bits["ID1"]) |>
  select(GROUP, intercept_ours, slope_ours, r_squared) |>
  left_join(article, by = "GROUP") |>
  mutate(GROUP = factor(GROUP, levels = c("pp", "ap", "cp"))) |>
  arrange(GROUP) |>
  mutate(slope_ratio_ours    = slope_ours / slope_ours[GROUP == "cp"],
         slope_ratio_article = slope_article / slope_article[GROUP == "cp"])

comparison |> mutate(across(where(is.numeric), \(x) round(x, 3)))
```

```{r}
#| label: compare-values
#| include: false
cmp <- function(g, col) comparison[[col]][comparison$GROUP == g]
```

## What differs

- **The patients' slope is steeper**: `r round(cmp("pp", "slope_ours"), 3)` s/bit against
  0.672 s/bit in the article, and the intercept becomes negative
  (`r round(cmp("pp", "intercept_ours"), 3)` s), which has no physical meaning: a movement
  cannot take a negative time at ID = 0.
- **The controls' slopes are a little steeper too**:
  `r round(cmp("ap", "slope_ours"), 3)` and `r round(cmp("cp", "slope_ours"), 3)` s/bit,
  against 0.118 and 0.097 s/bit.
- **Our patients are slower at ID2**: their mean movement time is
  `r round(mt_summary$mean[mt_summary$GROUP == "pp" & mt_summary$ID == "ID2"], 2)` s, with a
  very large spread between patients. This is why the patients' regression explains little
  of the variance ($R^2$ = `r round(cmp("pp", "r_squared"), 2)`): the difference between
  patients is much larger than the effect of ID.

