# Differences with the article (1.7)

## Putting our equations on the same scale as the article

In the regression section, ID was coded 1 for ID1 and 2 for ID2 (`ID_numeric`). The slopes of `regression_models` are therefore in seconds **per ID step**, and the intercepts are the movement times predicted at a step "0" that does not exist. The article fits Fitts' law on the index of difficulty in **bits** [@mottet2017; @fitts1954]. The two IDs of the task are
3.80 and 5.43 bits (x-axis of Fig 3 of the article), so one ID step is 1.63 bits.

To compare our equations with those of the article, we convert them to bits. The line goes through the same two group means; only the unit of the x-axis changes:

   - slope (s/bit) = slope (s/step) / 1.63;
   - intercept (s) = predicted MT at ID1 − slope (s/bit) × 3.80.
   

```{r}
#| label: compare-article
id_bits <- c(ID1 = 3.80, ID2 = 5.43)
step_bits <- unname(id_bits["ID2"] - id_bits["ID1"])

article <- tibble(
  GROUP = c("pp", "ap", "cp"),
  intercept_article = c(0.422, 0.622, 0.475),
  slope_article     = c(0.672, 0.118, 0.097)
)

comparison <- regression_models |>
  mutate(GROUP = as.character(GROUP),
         slope_ours     = slope / step_bits,
         intercept_ours = (intercept + slope * 1) - slope_ours * id_bits["ID1"]) |>
  select(GROUP, intercept_ours, slope_ours, r_squared) |>
  left_join(article, by = "GROUP") |>
  mutate(GROUP = factor(GROUP, levels = c("pp", "ap", "cp"))) |>
  arrange(GROUP) |>
  mutate(slope_ratio_ours    = slope_ours / slope_ours[GROUP == "cp"],
         slope_ratio_article = slope_article / slope_article[GROUP == "cp"])

comparison |> mutate(across(where(is.numeric), \(x) round(x, 3)))
```

```{r}
#| label: compare-values
#| include: false
cmp <- function(g, col) comparison[[col]][comparison$GROUP == g]
```

## What differs

- **The patients' slope is steeper**: `r round(cmp("pp", "slope_ours"), 3)` s/bit against 0.672 s/bit in the article, and the intercept becomes negative (`r round(cmp("pp", "intercept_ours"), 3)` s), which has no physical meaning: a movement cannot take a negative time at ID = 0.
- **The controls' slopes are a little steeper too**:
  `r round(cmp("ap", "slope_ours"), 3)` and `r round(cmp("cp", "slope_ours"), 3)` s/bit,
  against 0.118 and 0.097 s/bit.
- **Our patients are slower at ID2**: their mean movement time is
  `r round(mt_summary$mean[mt_summary$GROUP == "pp" & mt_summary$ID == "ID2"], 2)` s, with a very large spread between patients. This is why the patients' regression explains little of the variance ($R^2$ = `r round(cmp("pp", "r_squared"), 2)`): the difference between patients is much larger than the effect of ID.

These differences come from the simpler method used here. Several choices may matter (the
article does not describe all of them, so these are hypotheses):

1. **Data selected.** We removed ID3 and the third repetition to obtain a complete design.
   The article may have kept them, or handled the missing cells differently.
2. **How trials are averaged.** We used the mean of all the movements of each participant.
   Movement time is right-skewed (a few very slow corrective movements), so the mean is
   pulled up by the slowest movements, especially in the patients. A median, a log
   transformation or the removal of outlier trials would give smaller values and a smaller
   slope.
3. **Which ID.** We used the *nominal* ID of the task. Fitts' law is often fitted on the
   *effective* ID, computed from the actual spread of the endpoints, which can differ
   between groups.
4. **The full design.** The article also includes orientation and direction as factors
   (with Greenhouse–Geisser corrections), whereas we averaged over them.
5. **The regression model.** Our regression treats the two values of each participant as
   independent observations, while the article analyses them as repeated measures. This
   does not change the slope (the line goes through the two group means), but it makes the
   $R^2$ and the p-values of the regressions approximate.
   

## Are the conclusions the same?

```{r}
#| label: controls-only
# Do the two control groups differ? Same ANOVA without the patients.
mt_controls <- mt |> filter(GROUP != "pp") |> droplevels()
aov_controls <- ezANOVA(data = mt_controls, dv = MT, wid = SUBJ,
                        between = GROUP, within = ID, type = 3)$ANOVA
aov_controls
```

**Yes, the conclusions are the same**, because they rest on the *pattern* of the results,
which is identical, and not on the exact values:

- **Movement time increases with difficulty in all groups:** the three slopes are positive
  and the main effect of ID is significant.
- **Patients are much slower** (main effect of GROUP, the largest effect,
  $\eta^2_G$ = `r round(aov_tab$ges[aov_tab$Effect == "GROUP"], 2)`).
- **Patients are much more sensitive to accuracy constraints:** their slope is
  `r round(cmp("pp", "slope_ratio_ours"), 1)` times that of the young adults here, and
  `r round(cmp("pp", "slope_ratio_article"), 1)` times in the article. The ratio does not
  depend on how ID is coded, so it can be compared directly. The GROUP × ID interaction is
  significant (p < .001), so it passes even the stricter α = .01 used in the article.
